### 2015-03-16 bugfinding.tex
draft text for speeding up corral
initial text for bug finding section

%\textsc{Whoop} is currently working on an unmodified network device driver (the RealTek r8169 which is part of the official Linux kernel distribution). Our immediate future work includes applying invariant generation to tame the effect of our coarse over-approximation. We also plan to introduce counterexample feasibility checking to evaluate if a reported bug is real or spurious. Towards this, we plan to feed a counterexample generated by \textsc{Whoop} into a precise bug finder, such as Corral~\cite{lal2012corral}, which will be guided only to execution paths that may manifest the bug, to either confirm it or discard it as a false positive.

>>>>

\whoop is a sound but potentially imprecise static data race analyser. In our experience, developers appreciate soundness but do not want to get overwhelmed by countless false positive reports. Towards this we allow an external precise bug-finder to be plugged in the \whoop infrastructure. The race-related information generated by \whoop can then be used to accelerate the bug-finder a bug-finder for concurrent programs that can be easily plugged in our infrastructure. Currently we only support \corral, a state-of-the-art bug-finder used in Microsoft to analyse Windows device drivers. We chose \corral because it is open-source and because it accepts Boogie IVL, similar to \whoop. Our technique though should be able to accelerate an arbitrary bug-finder for concurrent programs.

Sound partial order reduction

Yield instrumentation

Granularities:

Full-force

Coarse

MemoryRegion

Speeding up Corral

%\textsc{Whoop} is currently working on an unmodified network device driver (the RealTek r8169 which is part of the official Linux kernel distribution). Our immediate future work includes applying invariant generation to tame the effect of our coarse over-approximation. We also plan to introduce counterexample feasibility checking to evaluate if a reported bug is real or spurious. Towards this, we plan to feed a counterexample generated by \textsc{Whoop} into a precise bug-finder, such as Corral~\cite{lal2012corral}, which will be guided only to execution paths that may manifest the bug, to either confirm it or discard it as a false positive.

### 2015-03-16 conclusion.tex
initial text for bug finding section

Compared to traditional data race detection tools that typically attempt to explore as many thread interleavings as feasible, and thus face scalability issues in the presence of realistic concurrent programs, \textsc{Whoop} not only entirely avoids reasoning about thread interleavings, but also allows the reuse of existing successful sequential verification techniques. The main limitation of our approach is that it can potentially report many false positives, as we over-approximate the device driver shared state. To tackle this problem, we use the race-related information that \whoop generates to speed up \corral, a state-of-the-art precise bug finder for concurrent programs.

>>>>

Compared to traditional data race detection tools that typically attempt to explore as many thread interleavings as feasible, and thus face scalability issues in the presence of realistic concurrent programs, \textsc{Whoop} not only entirely avoids reasoning about thread interleavings, but also allows the reuse of existing successful sequential verification techniques. The main limitation of our approach is that it can potentially report many false positives, as we over-approximate the device driver shared state. To tackle this problem, we use the race-related information that \whoop generates to speed up \corral, a state-of-the-art precise bug-finder for concurrent programs.

### 2015-03-16 intro.tex
initial text for bug finding section

In this paper, we present \whoop, a novel infrastructure for (i) automatically analysing the concurrent behaviour of Linux kernel modules written in the C language and (ii) accelerating external bug finders that can easily plug in our toolchain. \whoop is empowered by \emph{static pairwise lockset analysis}, a novel technique for automatically verifying the absence of data races in Linux kernel modules (e.g. device drivers and filesystems). The key idea behind our approach is that a module can be proven free from data races by (i) deriving a sound \emph{sequential} model that \emph{over-approximates} the originally concurrent program, (ii) instrumenting it for lockset analysis and race checking, and (iii) asserting that all accesses to the same shared resource are protected by at least one common lock. The immediate benefit is that our approach not only avoids reasoning about thread interleavings, and thus has the potential to scale well, but also allows the reuse of existing successful sequential verification techniques.

We then show that by exploiting the race-freedom guarantees that our static analysis provides, we can achieve a sound partial-order reduction that significantly accelerates \corral~\cite{lal2012solver}, a state-of-the-art bug finder that is used by Microsoft to analyse device drivers and other concurrent programs. Using \whoop and \corral we analysed 10 modules available in the Linux distribution.

To summarize, our contributions are as follows:
\begin{itemize}
\item We present \whoop, a practical tool for analysing the concurrent behaviour of Linux kernel modules and accelerating a plugged-in precise bug finder.
\item We discuss static pairwise lockset analysis,  a sound scalable technique for automatically verifying absence of data races in Linux kernel modules.
\item We show that we can achieve a sound partial-order reduction using our technique and significantly accelerate \corral, a state-of-the-art bug finder.
\item We analyse 10 Linux kernel modules, including various drivers and a filesystem.
\end{itemize}

>>>>

In this paper, we present \whoop, a novel infrastructure for (i) automatically analysing the concurrent behaviour of Linux kernel modules written in the C language and (ii) accelerating external bug-finders that can easily plug in our toolchain. \whoop is empowered by \emph{static pairwise lockset analysis}, a novel technique for automatically verifying the absence of data races in Linux kernel modules (e.g. device drivers and filesystems). The key idea behind our approach is that a module can be proven free from data races by (i) deriving a sound \emph{sequential} model that \emph{over-approximates} the originally concurrent program, (ii) instrumenting it for lockset analysis and race checking, and (iii) asserting that all accesses to the same shared resource are protected by at least one common lock. The immediate benefit is that our approach not only avoids reasoning about thread interleavings, and thus has the potential to scale well, but also allows the reuse of existing successful sequential verification techniques.

We then show that by exploiting the race-freedom guarantees that our static analysis provides, we can achieve a sound partial-order reduction that significantly accelerates \corral~\cite{lal2012solver}, a state-of-the-art bug-finder that is used by Microsoft to analyse device drivers and other concurrent programs. Using \whoop and \corral we analysed 10 modules available in the Linux distribution.

To summarize, our contributions are as follows:
\begin{itemize}
\item We present \whoop, a practical tool for analysing the concurrent behaviour of Linux kernel modules and accelerating a plugged-in precise bug-finder.
\item We discuss static pairwise lockset analysis,  a sound scalable technique for automatically verifying absence of data races in Linux kernel modules.
\item We show that we can achieve a sound partial-order reduction using our technique and significantly accelerate \corral, a state-of-the-art bug-finder.
\item We analyse 10 Linux kernel modules, including various drivers and a filesystem.
\end{itemize}

### 2015-03-16 whoop.tex
initial text for bug finding section

We have implemented our approach in \whoop, a prototype infrastructure that combines static lockset analysis with state-of-the-art compilation and sequential verification techniques to (i) soundly analyse Linux kernel modules for data races and (ii) accelerate a plugged-in bug finder.

@@

Figure~\ref{fig:whoop} depicts the \whoop infrastructure. Initially, \whoop accepts a kernel module written in C, together with a Linux environmental model. The environmental model (i.e. stub Linux header files) is required to ``close'' the module and allow it to be translated and subsequently analysed. Both the module and the environmental model are passed to the translation phase of \whoop (see Section~\ref{translation}), which outputs an over-approximated program written in the Boogie intermediate verification language (IVL)~\cite{deline2005boogiepl}. Next, \whoop performs its static lockset analysis (see Section~\ref{staticanalysis}) to detect all potential data races in the abstract program. Finally, \whoop uses information from the static analysis phase to accelerate bug finding with \corral (see Section~\ref{bugfinding}).

>>>>

We have implemented our approach in \whoop, a prototype infrastructure that combines static lockset analysis with state-of-the-art compilation and sequential verification techniques to (i) soundly analyse Linux kernel modules for data races and (ii) accelerate a plugged-in bug-finder.

@@

Figure~\ref{fig:whoop} depicts the \whoop infrastructure. Initially, \whoop accepts a kernel module written in C, together with a Linux environmental model. The environmental model (i.e. stub Linux header files) is required to ``close'' the module and allow it to be translated and subsequently analysed. Both the module and the environmental model are passed to the translation phase of \whoop (see Section~\ref{translation}), which outputs an over-approximated program written in the Boogie intermediate verification language (IVL)~\cite{deline2005boogiepl}. Next, \whoop performs its static lockset analysis (see Section~\ref{staticanalysis}) to detect all potential data races in the abstract program. Finally, \whoop uses information from the static analysis phase to accelerate bug-finding with \corral (see Section~\ref{bugfinding}).
